Global code motion of congruent instructions

### 2017-04-18 paper.tex
grammar

We have implemented the pass in \LLVM{}. It is based on Global Value Numbering
infrastructure available in \LLVM{}. The experimental results show an average
saving of 1\% on overall compile time on \SPEC{}. \GCM{} enables more inlining
and exposes more loop invariant code motion opportunities in majority of the
benchmarks. We have also seen execution time improvements in a few of \SPEC{}
benchmarks viz. mcf and sjeng, moreover, register spills reduced by 2\% on the
\SPEC{} benchmarks suite when compiled for \xlinux{}. This is an optimistic
algorithm in the sense that it considers all congruent instructions in a
function as potential candidates. We make an extra effort to hoist candidates by
partitioning the potential candidates in a way to enable partial hoisting in
case common hoisting points for all the candidates cannot be found. We also
formalize how register pressure will reduce as a result of \gcm{}, and how
\gcm{} increases instruction level parallelism thereby enabling more \ooo{}
execution on modern architectures.
\end{abstract}

@@

We see some nice gains in code size at -Os with \gcm{} (403.gcc, 447.dealII, and
483.xalancbmk) as shown in Table~\ref{tab:code-size}.  The code size listed here
is the text size delta (from linux size command) of the final executable for
each benchmark.  Some benchmarks like 473.astar increased in code size by almost
2\%. This is because once the code-size of a function decreases (due to \GCM{}),
it becomes cheaper to inline. As we can see in Table~\ref{tab:stats}, four more
functions got inlined in 473.astar. Except for 447.dealII, 450.soplex,
482.sphinx3 and 483.xalancbmk all other benchmarks got more functions
inlined. Also, since the pass runs early, it affects many optimizations which
rely on the number of instructions, length of the use-def chain, and other
metrics.

>>>>

We have implemented the pass in \LLVM{}. It is based on Global Value Numbering
infrastructure available in \LLVM{}. The experimental results show an average
saving of 1\% on overall compile time on \SPEC{}. \GCM{} enables more inlining
and exposes more loop invariant code motion opportunities in majority of the
benchmarks. We have also seen execution time improvements in a few of \SPEC{}
benchmarks viz. mcf and sjeng, moreover, register spills reduced by 2\% on the
\SPEC{} benchmarks suite when compiled for \xlinux{}. This is an optimistic
algorithm in the sense that it considers all congruent instructions in a
function as potential candidates. We make an extra effort to hoist candidates by
partitioning the potential candidates in a way to enable partial hoisting in
case a common hoisting point for all the candidates cannot be found. We also
formalize how register pressure will reduce as a result of \gcm{}, and how
\gcm{} increases instruction level parallelism thereby enabling more \ooo{}
execution on modern architectures.
\end{abstract}

@@

We see some nice gains in code size at -Os with \gcm{} (403.gcc, 447.dealII, and
483.xalancbmk) as shown in Table~\ref{tab:code-size}.  The code size listed here
is the text size delta (from linux size command) of the final executable for
each benchmark.  On the other hand, some benchmarks like 473.astar increased in
code size by almost 2\%. This is because once the code-size of a function
decreases (due to \GCM{}), it becomes cheaper to inline. As we can see in
Table~\ref{tab:stats}, four more functions got inlined in 473.astar. Except for
447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got
more functions inlined. Also, since the pass runs early, it affects many
optimizations which rely on the number of instructions, length of the use-def
chain, and other metrics.
